---
title: "Scottish temperatures"
subtitle: "Explore trends in minimum and maximum temperatures across Scotland by month."
execute: 
  echo: false
  eval: true
  warning: false
  message: false
  freeze: auto
---

```{r}
# Packages
library(tidyverse)
library(showtext)
library(ggtext)
library(nrBrand)
library(glue)
library(ggview)
library(rnaturalearth)
library(ggiraph)
library(htmltools)
library(cowplot)
library(grid)
```

```{r}
# Data
weather <- read_csv("data/weather.csv")
metadata <- read_csv("data/metadata.csv")
uk <- ne_states(country = "united kingdom")
scot_map <- uk |>
  filter(geonunit == "Scotland")
```

```{r}
# Fonts
font_add_google("Oswald")
font_add_google("Nunito")
showtext_auto()
showtext_opts(dpi = 300)
```

```{r}
# Parameters
title_font <- "Oswald"
body_font <- "Nunito"
bg_col <- "#F2F4F8"
text_col <- "#151C28"
```

```{r}
# Settings
knitr::opts_chunk$set(dev.args = list(bg = bg_col))
```

```{r}
# Data wrangling
station_levels <- metadata |>
  arrange(desc(lat)) |>
  pull(station)

plot_data <- weather |>
  select(station, year, month, min_temp, max_temp) |>
  mutate(
    month = factor(month.name[month], levels = month.name),
    station = factor(station, levels = station_levels)
  ) |>
  drop_na()

max_data <- plot_data |>
  group_by(station, month) |>
  slice_max(max_temp) |>
  ungroup() |>
  select(station, month, year, max_temp)

min_data <- plot_data |>
  group_by(station, month) |>
  slice_min(min_temp) |>
  ungroup() |>
  select(station, month, year, min_temp)

min_max_data <- plot_data |>
  group_by(
    station, month
  ) |>
  summarise(
    min_year = min(year),
    max_year = max(year),
    min_temp = min(min_temp, na.rm = TRUE),
    max_temp = max(max_temp, na.rm = TRUE)
  ) |>
  ungroup()
```

The Met Office makes monthly data available from a selection of long-running historic weather stations. The series typically range from 50 to more than 100 years in length. There are 8 stations in Scotland which are still actively collecting data. The data collected includes mean daily minimum and maximum temperatures.

::: {.grid}

::: {.g-col-12 .g-col-md-6}

```{r}
#| file: R/map.R
#| fig-asp: 1
#| fig-alt: "Map of Scotland showing locations of 8 weather stations."
```

:::

::: {.g-col-12 .g-col-md-6}

```{r}
#| file: R/legend.R
#| fig-asp: 1
#| fig-alt: "Mini chart for Leuchars station with annotations explaining how to interpret it."
```


:::

:::

<br>

```{r}
# Dropdown
options_vec <- month.name
names(options_vec) <- month.name
dropdown_html <- tags$div(
  style = "text-align: center; margin-bottom: 10px;",
  tags$select(
    id = "dropdown",
    lapply(names(options_vec), function(name) {
      tags$option(value = options_vec[[name]], name)
    })
  )
)
browsable(
  tagList(
    dropdown_html
  )
)
```

```{css}
#dropdown {
  border-radius: 4px;
  border: 2px solid #AAAAAA;
  padding-bottom: 2px;
  width: 100%;
}

.grid {
  gap: 0px !important;
}
```

Average minimum and maximum daily temperatures (°C) per month across eight different weather stations in Scotland.

```{r}
#| file: R/chart.R
```

```{r}
#| fig-asp: 0.9
#| fig-alt: "Small multiple charts showing monthly minimum and maximum temperature at each weather station in Scotland."
g_int <- girafe(
  ggobj = g_chart,
  options = list(
    opts_hover(css = ""),
    opts_toolbar(hidden = c("saveaspng", "fullscreen"), saveaspng = FALSE)
  )
)
browsable(
  tagList(
    g_int,
    tags$script(HTML("
      function filterByDropdown() {
        const selected = document.getElementById('dropdown').value;
        document.querySelectorAll('[data-id]').forEach(el => {
          if (el.getAttribute('data-id') === selected) {
            el.setAttribute('opacity', '1');
            el.classList.add('selected');
          } else {
            el.setAttribute('opacity', '0');
            el.classList.remove('selected');
          }
        });
      }

      document.getElementById('dropdown').addEventListener('change', filterByDropdown);
      document.addEventListener('DOMContentLoaded', filterByDropdown);
    "))
  )
)
```

<br>

**Note**: Only active stations are included, which means that Nairn (closed 2014) and Paisley (closed 2011) are excluded. Braemar station had a slight change in location in April 2005.

**Data**: [Historic station data](https://www.metoffice.gov.uk/research/climate/maps-and-data/historic-station-data) from the Met Office. Accessed 10 July 2026.

**Graphic**: {{< fa brands linkedin >}} [nicola-rennie](https://www.linkedin.com/in/nicola-rennie/) {{< fa brands github >}} [nrennie](https://github.com/nrennie) {{< fa globe >}} [nrennie.rbind.io](https://nrennie.rbind.io/)
